---
title: "using-reclanc"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{using-reclanc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(reclanc)
library(SummarizedExperiment)
library(parsnip)
```

# Fitting

To create new centroids from existing expression data, use the `clanc` function.

`reclanc` provides some synthetic expression data that we can work with:

```{r}
lapply(synthetic_expression, head)
```

These data include 12 samples, 6 of class "A" and 6 of class "B", with 100 genes.

`reclanc` is able to ingest a variety of common formats for expression data. Objects like `SummarizedExperiments` and `ExpressionSets` are frequently used in bioinformatic analyses and arrange their data so samples are in columns and genes are in rows. This is in conflict with the expected formula input in base R, where predictors (genes) and outcomes (classes) are in columns. `reclanc` eases this friction by expecting input in its most common format, abstracting away the wrangling aspect of analysis.

Because of this, there are two broad categories of input - 'wide data' and 'tall data'.


## Wide inputs

Wide inputs require data that has both predictors and outcomes as columns, together, in a single `data.frame`.

### Formula

```{r}
form_data <- cbind(
  class = synthetic_expression$classes,
  as.data.frame(t(synthetic_expression$expression))
)
head(form_data[1:5])
```

```{r}
clanc(class ~ ., form_data, active = 5)
```

### `recipe`

`reclanc` also supports `tidymodels` workflows:

```{r}
discrim_linear() |>
  set_engine("clanc", active = 5) |>
  fit(class ~ ., data = form_data)
```

## Tall inputs

Tall inputs require genes as rows and samples as columns

### `data.frame`/`matrix`

It is often convenient to supply a `data.frame`, particularly after some data-munging has been done. Both `data.frame` and `matrix` inputs require expression with genes as column names and sample IDs as rownames, as well as a factor vector of classes:


```{r}
clanc(
  synthetic_expression$expression,
  classes = synthetic_expression$classes,
  active = 5
)
```

### `SummarizedExperiment`

Some common formats for expression are `SummarizedExperiment`s and `ExpressionSet`s:

```{r}
se <- SummarizedExperiment(
  synthetic_expression$expression,
  colData = DataFrame(class = synthetic_expression$classes)
)
se
```

We can specify the name of the `colData` (`pData` for `ExpressionSet`s) column that contains the classes with the `classes` argument:

```{r}
fit <- clanc(
  se,
  classes = "class",
  active = 20,
  assay = 1 # Index of assay - SummarizedExperiments only
)
fit
```

# Predicting

The fit can then be used to predict classes of new samples from new data. The new data can come in the form of a `matrix`, `data.frame`, `SummarizedExperiment`, or `ExpressionSet`, with the same expected input


Using `type = "class"` will predict classes using the metric provided by Alan Dabney in the original ClaNC paper.

```{r}
predict(fit, new_data = se, type = "class")
```

However, particularly if comparing across datasets that may have been transformed differently, it may be more accurate to use a correlation based metric:

```{r}
predict(fit, new_data = se, type = "numeric", method = "spearman")
```
